Near instantaneous response of mesophyll conductance to light availability allows shade leaves to lie in wait

### 2015-08-02 manuscript/manuscript_wtc3.Rmd
added discussion to manuscript
made pngs scripts and figures for esa talk

## Stomatal conductance, mesophyll conductance and CO~2~ drawdown
Mean g~s~ was significantly higher in shade compared to sun leaves (+18%), under their local light environment (P = 0.004, Table2).  Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001, Figure `r gebarfig`a). In sun leaves, g~s~ and A showed optimal behavior *as in (cite)* with diminishing rates of A with highest values of g~s~ (Figure `r agmgsfig`a).  However, in shade leaves this relationship was clearly uncoupled.  With increased light availability, both increases in A and g~s~ in shade leaves extened the relationship of A and g~s~ seen in sun leaves. 
\
The observed carbob isotope discrimination was postively correlated with C~i~:C~a~ for all leaf types (P < 0.001) with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light (Figure `r deltafig`). Here, the carbon isotope discrimination associtaed with g~m~ accounted for the majority of $\Delta$~o~ (69.7%pm;0.4 %) and varied little across measurement temperatures, leaf types, or warming treatments. Mean g~m~ was significantly higher in sun compared to shade leaves (+27%), under their local light environment (P < 0.001, Table2).  Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001). Relationships between g~m~ and A were nearly proportional within leaf types and light treatments (Figure `r gebarfig`b,c). Photosynthesis scaled postively with increases in g~m~ for all leaves and the large increases in g~m~ in shade leaves in high light resulted in the highest rates of A (Figure `r agmgsfig`b). No differences in g~s~ nor g~m~ were detected with the warming treatment within leaf types or light treatments. Additionally, neither g~s~ nor g~m~ varied significantly across measurements campaigns (Figures `r get1fig`b and `r get2fig`b, respectively) and a slightly negative relationship with g~m~ and increasing leaf temperature.

\
The combination of a non-limiting supply of leaf K and more open stomata lead to a significant increase in C~i~ in shade compared to sun leaves (Figure `r ciccfig`a). As increases in A in shade leaves were associated with higher rates of g~s~ the C~i~ was similar than under low light but still greater than sun leaves. As the drop in CO~2~ for intecellular spaces to the cholorplast, C~i~-C~c~, measures the coordiation between gm and A [@caemmerer2014temperature] an increase in this drawdown was detected in shade leaves at high light (Figure `r ciccfig`b,c).  This was a direct result of the increase in g~m~ and its subsequent direct effect on A, when compared to both sun and shade leaves that are acclimated to their local light environment. CO~2~ drawdown from C~a~ to C~i~ and C~i~ to C~c~ were both relatively stable acorss the range of temperatures measured and gas exchange campaigns (Figure `r get2fig`c and `r get1fig`c, respectively).

# Discussion

photosynthetic capacity: Mesophyll conductance asymmetrically affects key photosynthetic parameters and an assumption of infinite mesophyll conductance leads to underestimation of the maximum carboxylation rate Vcmax, maximum electron transport rate Jmax [@sun2014asymmetrical].

>>>>

## Stomatal conductance, mesophyll conductance and CO~2~ drawdown
Mean g~s~ was significantly higher in shade compared to sun leaves (+18%), under their local light environment (P = 0.004, Table2).  Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001, Figure `r gebarfig`a). In sun leaves, g~s~ and A exhibited optimal behavior with diminishing rates of A with highest values of g~s~ (Figure `r agmgsfig`a).  However, in shade leaves this relationship was clearly uncoupled.  With increased light availability, both increases in A and g~s~ in shade leaves extened the relationship of A and g~s~ seen in sun leaves. 
\
The observed carbob isotope discrimination was postively correlated with C~i~:C~a~ for all leaf types (P < 0.001) with larger $\Delta$~o~ detected for sun leaves and shade leaves at high light (Figure `r deltafig`). Here, the carbon isotope discrimination associtaed with g~m~ accounted for the majority of $\Delta$~o~ (69.7%pm;0.4 %) and varied little across measurement temperatures, leaf types, or warming treatments. Mean g~m~ was significantly higher in sun compared to shade leaves (+27%), under their local light environment (P < 0.001, Table2).  Following an increase in available light, g~s~ of shade leaves was significantly greater than both shade leaves at low light and sun leaves (P < 0.001). Relationships between g~m~ and A were nearly proportional within leaf types and light treatments (Figure `r gebarfig`b,c). Photosynthesis scaled postively with increases in g~m~ for all leaves and the large increases in g~m~ in shade leaves in high light resulted in the highest rates of A (Figure `r agmgsfig`b). No differences in g~s~ nor g~m~ were detected with the warming treatment within leaf types or light treatments. Additionally, neither g~s~ nor g~m~ varied significantly across measurements campaigns (Figures `r get1fig`b and `r get2fig`b, respectively) and a slightly negative relationship with g~m~ and increasing leaf temperature.

\
The combination of a non-limiting supply of leaf K and more open stomata lead to a significant increase in C~i~ in shade compared to sun leaves (Figure `r ciccfig`a). As increases in A in shade leaves were associated with higher rates of g~s~ the C~i~ was similar than under low light but still greater than sun leaves. As the drop in CO~2~ for intecellular spaces to the cholorplast, C~i~-C~c~, measures the coordiation between g~m~ and A [@caemmerer2014temperature] an increase in this drawdown was detected in shade leaves at high light (Figure `r ciccfig`b,c).  This was a direct result of the increase in g~m~ and its subsequent direct effect on A, when compared to both sun and shade leaves that are acclimated to their local light environment. CO~2~ drawdown from C~a~ to C~i~ and C~i~ to C~c~ were both relatively stable acorss the range of temperatures measured and gas exchange campaigns (Figure `r get2fig`c and `r get1fig`c, respectively).

# Discussion
Here we show that dynamic physiological responses in shade leaves to altered light environments has important implications for upscaling leaf level measurements to the canopy. Tree canopies may maintain the ability to upregulate A beyond the constraints of their local light environment, even at the cost of valuable resources. Empirical data accounting for the total conductance of CO2 (both g~s~ and g~m~) reveal that these responses can lead to leaf C gain equivalent to sun leaves and may happen quicker than previously thought. Thus, failure to integrate the effects of g~m~ into process based tree growth models may underestimate the ability of canopies to alter their photosynthetic capacity when environmental conditions change. Additionally, with measurements recorded across a large natural range of temperatures we were unable to detect any effects of a +3 &deg;C warming treatment on leaf physiology.

## sub heading
The allocation of N constrains photosynthesis rates and is thus a key trait in determining the relative contribution of individual leaves to canopy C gain.  Decreasing light availability should decrease the investment into photosynthetic enzyme within a canopy (Mooney and Gulmon 1979).  As a result, acclimation of photosynthetic capacity to irradiance is typically reflected in the key photosynthetic biochemical parameters Vcmax and Jmax [@farquhar1980biochemical]. Our data agree with these conventional conclusions as the distribution of N, both biochemical parameters and A were all reduced in shade leaves.

\
Photosynthesis in trees is also limited by the ability to supply water to the upper canopy. Using a two-leaf model Peltoniemi et al. [-@peltoniemi2012co] theorizes that optimal N distribution will be proportional to light distribution only if leaf K is also optimally distributed. In this study, variation in leaf N distribution and A rates were not associated with subsequent changes in leaf K between sun and shade leaves. This is meaningful as foliage tends to operate at greater water limitations in the upper canopy, despite the ability of sun leaves to develop more effective vascular systems [@sellin2008effects;@niinemets2012optimization]. As both sun and shade leaves were acclimated to their local light environment this sub-optimal distribution of water supply has the potential to diminish the C return for the whole canopy.

\
The inefficient distribution of resources led to decreased instantaneous WUE in shade leaves throughout the experiment.  Additionally, consistently higher leaf $\delta$^13^C in shade leaves suggests that this pattern was constant across leaf lifespans.  From a canopy perspective this pattern in WUE appears to be detrimental to C gain as A in sun leaves was characterized by low rates of g~s~ and low C~i~ compared to similar Eucalyptus species [see @drake2014capacity].  As whole canopy C gain integrates the efficiency of all leaves, this begs the question of why shade leaves maintained a lower WUE while the photosynthetic efficiency of sun leaves appears at least partially constrained by water supply.

\
The pattern of wasteful water use in shade leaves is important as we hypothesized that gs and A would be proportional across sun and shade leaves.  In sun leaves variation in A and g~s~ were strongly correlated, exhibiting behaviour agreeing with optimal stomatal theory. However, lower rates of A in shade leaves were not coupled with decreases in g~s~ which resulted in the observed decreases in WUE. This is significant as optimal stomatal behaviour has been reported across a wide range of ecosystems and plant functional types; however, empirical data is often collected only on sun leaves [see @prentice2014balancing; @lin2015optimal].  As a result, the often used economic framework of balancing costs of using water versus N allocation to predict photosynthesis [@wright2003least] may break down when considering all leaves within a tree canopy.

\
This behaviour may be explained by the response of shade leaves to increases in light availability. Contrary to hypothesis, shade leaves were able to increase A to rates equivalent or even outperforming sun leaves under identical light intensity. It is possible that sustaining higher g~s~, at the cost of larger transpiration losses, is a strategy to increase the utilization efficiency of sunflecks. These results should not be entirely unexpected as shade leaves experience interchanging periods of sun and shade, despite canopy light extinction [@pearcy1990sunflecks]. Additionally, empirical studies that assess photosynthetic responses to sunflecks generally focus on seedlings [@kuppers1993leaf;@leakey2002relative] and understory plants [@chazdon1991importance; @brantley2009contribution]. Our findings highlight a critical need for empirical measurements of shade leaves under dynamic light environments in order to accurately scale C gain from leaf to canopy.

\
Increasing the drawdown of CO~2~ into the intercellular air spaces via g~s~, however, does not necessarily infer concomitant increases in A. This is because CO~2~ supply to the site of carboxylation, via g~m~, may be anatomically constrained. In this study A and g~m~ scaled positively across leaf types and increased proportionately with increased light intensity in shade leaves.  Research has suggested that aquaporins can facilitate increases in the CO~2~ permeability of the cell membranes resulting in rapid modulation of mesophyll conductance [@hanba2004overexpression; @heinen2009role; @li2014plant]. This provides a potential explanation for our observed rapid increases in g~m~ but is untested in leaves of woody species. Our findings support a growing wealth of evidence that g~m~ is highly variable and can respond to environmental variables [@flexas2008mesophyll]. Here we provide empirical data showing g~m~ not only varies within a canopy but the up-regulation of g~m~ plays a critical role in the photosynthetic response of shade leaves to sunflecks.

\
Similar to Laisk [-@laisk2005adjustment] we show evidence that optimal acclimation, in this case physiological adjustment of the drawdown of CO~2~ in shade leaves, is possibly directed towards occasional maximums of light availability over adaptations towards a low light environment.  If shade leaves ‘lie in wait’ for sunflecks then perhaps we should consider an alternate leaf economic strategy to maximise C gain, beyond conventional trade-offs associated with canopy resource distribution.  If the valuation of sunflecks as a C resource is large enough then costs of sub-optimal stomatal behaviour in shade leaves could be offset over the leaf lifespan.  The potential benefits of C gain in leaves constrained by low light to levels equivalent to sun leaves may be large enough to tolerate excess losses from transpiration.  However, accounting for the heterogeneous nature of light within a canopy remains a current challenge for empirical and modelling studies. Thus, models which predict leaf photosynthesis from N distribution within a canopy will be flawed unless inclusion of canopy light extinction and the integrations of sunflecks on shade leaves are included [@de1997simple].

\
Furthermore, our findings have important implications for process based models that predict canopy C gain from rates of leaf photosynthesis. Although the distributions of resources, such as N and water, define leaf photosynthetic capacity it is the physiological behaviour of individual leaves which actually determine C gain. The observed physiological behaviour leaves in the shade, at both low and high irradiance, suggests that current theories of leaf optimal behaviour are incomplete.  Furthermore, gm must now be considered as a dynamic process in modelling whole canopy C gain.  Additional empirical data are needed to determine both the mechanisms responsible for and the capacity of gm to rapidly increase CO2 drawdown. To improve our ability to predict whole canopy C gain future research should prioritize the incorporation of both sun and shade leaf physiology, which may be optimized differently.
